Supplementary Materials

### 2023-07-23 manuscript/manuscript.tex
ms updates from italy

(see Appendix Equation \ref{eq:Qinf}). This general equation can accommodate
models of polygenic selection as long as the equilibrium additive fitness
variation $V_{A,i}$ can be specified and the change in variance due to
selection can be approximated as a geometric decay, i.e. $\Delta V_{A,i} =
-\kappa V_{A,i}$ \parencite{Bulmer1971-ae,Keightley1988-eq,Walsh2018-bt}. This
is usually a reasonable assumption since within-generation selection removes a
fraction of phenotypic variation from the population, and some fraction of that
is additive genetic variation \parencite{Bulmer1971-ae,Keightley1988-eq}.

@@

Our PhastCons Priority model fits marginally better than the CADD 6\% model
consistently across all three populations (Supplementary Table
\ref{supp:tbl-r2}). This is likely because the PhastCons Priority model has
more DFE parameters, and thus can fit finer-scale differences among the four
feature types. By contrast, the CADD 6\% track must fit patterns of diversity
to conserved sites under a single DFE. The CADD 8\% model fit slightly worse
than the CADD 6\% across all populations, and the Feature Priority model had
the poorest-fit.

@@

We find across all populations, our mutation rate estimates from CADD-based
models are roughly consistent with pedigree-based estimates (Figure
\ref{fig:figure-4}A), consistent with recent work \parencite{Murphy2022-sj}.
Our full-track CADD 6\% model estimates the mutation rate as $\widehat{\mu} =
1.56 \times 10^{-8}$ for YRI, $1.64 \times 10^{-8}$ for CEU, and $1.60 \times
10^{-8}$ for CHB reference samples (Supplementary Table \ref{supp:tbl-r2}). As
expected, the sparse-track CADD model mutation rate estimates are nearly
identical between the B' and classic BGS methods (Figure \ref{fig:figure-4}A
top row).

@@

Second, we noticed that across all populations and sparse/full tracks, the CADD
6\% model consistently led to slightly higher mutation rates than the CADD 8\%
model (Figure \ref{fig:figure-4}A bottom row; Supplementary Table
\ref{supp:tbl-r2}). This same pattern was observed in Murphy et al. too (2023;
Appendix 1, Figure 16). This behavior suggests a non-identifiability issue
between a slightly higher per-basepair mutation rate and annotation tracks that
contain more conserved sequence. This is expected from theory, since both
classic BGS and SC16 models only depend on mutation rate through the compound
parameter $\mu L$, where $L$ is the length of the conserved segment. Even
though our method is much more robust to inclusion of non-conserved regions
like introns, we still observe this non-identifiability issue.

>>>>

(see Supplementary Materials Equation \ref{supp-eq:Qinf}). This general equation can accommodate
models of polygenic selection as long as the equilibrium additive fitness
variation $V_{A,i}$ can be specified and the change in variance due to
selection can be approximated as a geometric decay, i.e. $\Delta V_{A,i} =
-\kappa V_{A,i}$ \parencite{Bulmer1971-ae,Keightley1988-eq,Walsh2018-bt}. This
is usually a reasonable assumption since within-generation selection removes a
fraction of phenotypic variation from the population, and some fraction of that
is additive genetic variation \parencite{Bulmer1971-ae,Keightley1988-eq}.

@@

Our PhastCons Priority model fits marginally better than the CADD 6\% model
consistently across all three populations (Supplementary Materials Table
\ref{supp:tbl-r2}). This is likely because the PhastCons Priority model has
more DFE parameters, and thus can fit finer-scale differences among the four
feature types. By contrast, the CADD 6\% track must fit patterns of diversity
to conserved sites under a single DFE. The CADD 8\% model fit slightly worse
than the CADD 6\% across all populations, and the Feature Priority model had
the poorest-fit.

@@

We find across all populations, our mutation rate estimates from CADD-based
models are roughly consistent with pedigree-based estimates (Figure
\ref{fig:figure-4}A), consistent with recent work \parencite{Murphy2022-sj}.
Our full-track CADD 6\% model estimates the mutation rate as $\widehat{\mu} =
1.56 \times 10^{-8}$ for YRI, $1.64 \times 10^{-8}$ for CEU, and $1.60 \times
10^{-8}$ for CHB reference samples (Supplementary Materials Table
\ref{supp:tbl-r2}). As expected, the sparse-track CADD model mutation rate
estimates are nearly identical between the B' and classic BGS methods (Figure
\ref{fig:figure-4}A top row).

@@

Second, we noticed that across all populations and sparse/full tracks, the CADD
6\% model consistently led to slightly higher mutation rates than the CADD 8\%
model (Figure \ref{fig:figure-4}A bottom row; Supplementary Materials Table
\ref{supp:tbl-r2}). This same pattern was observed in Murphy et al. too (2023;
Appendix 1, Figure 16). This behavior suggests a non-identifiability issue
between a slightly higher per-basepair mutation rate and annotation tracks that
contain more conserved sequence. This is expected from theory, since both
classic BGS and SC16 models only depend on mutation rate through the compound
parameter $\mu L$, where $L$ is the length of the conserved segment. Even
though our method is much more robust to inclusion of non-conserved regions
like introns, we still observe this non-identifiability issue.

### 2023-07-23 manuscript/supplementary.tex
ms updates from italy

\begin{align}
  \label{eq:Qinf}
  Q_\infty &= \frac{1}{\kappa + r(1-\kappa)}.
\end{align}

@@

\subsection{Table of $R^2$ Values}

\begin{table}
    \label{supp:tbl-r2}
\centering

@@

\begin{table}
    \label{supp:tbl-r2-strongsel}
\centering

@@

\begin{figure}[htbp]
  \label{suppfig:jackknife-iqr}
  \centering
  \includegraphics[width=0.7\textwidth]{figures/supplementary/iqr_jackknife_trim.pdf}

  \caption{The distribution of block jackknife samples for the Feature Priority
  model fit on YRI data. The ranges at the top of the figure indicate the
untrimmed and the 2 IQR-trimmed jackknife intervals. Trimming off the numeric
optimization outliers leads to more realistic estimates of uncertainty.}

\end{figure}

We use a block jackknife procedure to calculate the uncertainty of our
estimates. This approach is unfortunately computationally intensive, since a
10Mbp block width requires $\approx 280$ blocks to cover 2.8Gbp of autosomal
sequence and each fit requires thousands of random starts to reliably find
consistent optima. We use a parallelization strategy where a set number of
evenly-spaced positions along the genome that are chosen to approximately cover
all 2.8Gbps of autosomal sequence. Each of these positions is sent out across a
computing cluster, which excludes the 10Mbp block there and fits on the
remaining data. Then, using these jackknife samples we calculated the jackknife
standard error for each of parameters \parencite{Wasserman2004-ip}. In some
cases, our jackknife standard error estimates were very large. Upon inspection
of optimization diagnostic plots (see Supplementary Materials Section
\ref{suppfig:diag-plot}) and histograms of jackknife fits (Supplementary Figure
\ref{suppfig:jackknife-iqr}), these were due to rare aberrant optima. We
removed these using a trimmed-interquartile range, which removes points that
are 2 interquartile ranges below the first quartile or above the third quartile
(1.5 is standard, but we chose to be more conservative to not underestimate
uncertainty).

@@

\begin{figure}[htbp]
    \label{suppfig:rescale}
  \centering
  \includegraphics[]{figures/supplementary/rescale_pi_pred_plot.pdf}

  \caption{Chromosome 2 predictions on the YRI samples, with locally-rescaled
      model fits. Observed data is the dark gray line, and the normal MLE for
      the PhastCons Priority model is the blue line. The locally-rescaled
      predictions are the green line. The dashed red line are the prediction
      using the standard B' map (without local-rescaled) and the maximum likelihood
      estimates from the locally-rescaled fits. The large discrepancy in this shows
      that estimates are highly dependent on the B' map.}
\end{figure}

\begin{figure}[htbp]
    \label{suppfig:rescale-b}
  \centering
  \includegraphics[]{figures/supplementary/rescale_Bs.pdf}

  \caption{The standard (dark gray) and locally-rescaled (blue) B' maps for different
      $\mu = 1.6 \times 10^{-8}$ and three different selection coefficients.
      For $s=10^{-5}$ ($2N_e s = 0.2$), locally-rescaling alters the predicted
      reduction so that it is essentially insignificant ($B \approx 1$). For
  mid-strength selection $s=10^{-4}$ ($2N_e s = 2$), there is only a very
slight difference between standard and locally-rescaled B' maps. Finally, for
strong selection $s=10^{-3}$ ($2N_e s = 20$), locally-rescaling does not change
the B' maps, as expected. }
\end{figure}

\begin{figure}[htbp]
    \label{suppfig:rescale-dfe}
  \centering
  \includegraphics[]{figures/supplementary/rescale_DFE.pdf}

@@

\end{figure}

@@

\begin{figure}[htbp]
    \label{suppfig:strong-sel-dfe}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/supplementary/strongsel_grid_dfe.pdf}
  \caption{The DFE estimates for the strong selection grid (up to $s=10^{-1}$).}
\end{figure}

\begin{figure}[htbp]
    \label{suppfig:sel-ident}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/supplementary/ave_sel_vs_pi0.pdf}

@@

\end{figure}

\begin{figure}[htbp]
    \label{suppfig:strong-sel-dfe-pop-phastcons}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/supplementary/feature_priority_pop_dfe_strongselgrid.pdf}

  \caption{The DFE estimates for the strong selection grid (up to $s=10^{-1}$)
  under the Feature Priority model, across populations.}
\end{figure}

\begin{figure}[htbp]
    \label{suppfig:strong-sel-dfe-pop-feature}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/supplementary/phastcons_priority_pop_dfe_strongselgrid.pdf}

  \caption{The DFE estimates for the strong selection grid (up to $s=10^{-1}$)
  under the PhastCons Priority model, across populations.}
\end{figure}

@@

\begin{figure}[htbp]
    \label{suppfig:sim-exp}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/supplementary/subfigure_sim_mle_expansion.pdf}

@@

\end{figure}

\begin{figure}[htbp]
  \label{suppfig:sim-rec}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/supplementary/subfigure_sim_mle_recessive.pdf}

@@

\end{figure}

@@

\begin{figure}[htbp]
  \label{suppfig:resid-cadd}
  \centering
  \includegraphics[width=0.6\textwidth]{figures/supplementary/cadd_residual.pdf}
  \caption{Residuals from the CADD 6\% sparse track plotted against the fraction of
  basepairs in a window annotated by a CADD region.}
\end{figure}

\begin{figure}[htbp]
  \label{suppfig:resid-sweeps}
  \centering
  \includegraphics[width=0.6\textwidth]{figures/supplementary/sweep_residuals.pdf}

@@

\end{figure}

\begin{figure}[htbp]
  \label{suppfig:resid-b}
  \centering
  \includegraphics[width=0.6\textwidth]{figures/supplementary/b_resid.pdf}
  \caption{Residuals from the CADD 6\% model plotted against the predicted $B(x)$
           value.}
\end{figure}

\begin{figure}[htbp]
  \label{suppfig:resid-pc-fit}
  \centering
  \includegraphics[width=0.6\textwidth]{figures/supplementary/phastcons_resid_lof.pdf}
  \caption{Residuals from the PhastCons Priority model plotted against the
  average LoF selection coefficient across genes in a megabase window
(estimated from \cite{Agarwal2023-un}). These results show that the remaining likely 
selection signal is not dependent on the model used (PhastCons Priority versus CADD 6\%).}
\end{figure}

@@

\begin{figure}[htbp]
  \label{suppfig:dfe-ceu}
  \centering
  \includegraphics[width=0.9\textwidth]{figures/supplementary/figure_ceu_dfe.pdf}
  \caption{The DFE across models for the CEU reference sample (analogous to the
  DFE for YRI samples in Figure \ref{fig:figure-3}.}
\end{figure}

\begin{figure}[htbp]
    \label{suppfig:dfe-chb}
  \centering
  \includegraphics[width=0.9\textwidth]{figures/supplementary/figure_chb_dfe.pdf}
  \caption{The DFE across models for the CHB reference sample (analogous to the
  DFE for YRI samples in Figure \ref{fig:figure-3}.}
\end{figure}

>>>>

\begin{align}
  \label{supp-eq:Qinf}
  Q_\infty &= \frac{1}{\kappa + r(1-\kappa)}.
\end{align}

@@

\subsection{Table of $R^2$ Values}
\label{supp:tbl-r2}
\begin{table}
\centering

@@

\subsection{Table of $R^2$ Values (Strong Selection Grid)}
\label{supp:tbl-r2-strongsel}
\begin{table}
\centering

@@

We use a block jackknife procedure to calculate the uncertainty of our
estimates. This approach is unfortunately computationally intensive, since a
10Mbp block width requires $\approx 280$ blocks to cover 2.8Gbp of autosomal
sequence and each fit requires thousands of random starts to reliably find
consistent optima. We use a parallelization strategy where a set number of
evenly-spaced positions along the genome that are chosen to approximately cover
all 2.8Gbps of autosomal sequence. Each of these positions is sent out across a
computing cluster, which excludes the 10Mbp block there and fits on the
remaining data. Then, using these jackknife samples we calculated the jackknife
standard error for each of parameters \parencite{Wasserman2004-ip}.

\begin{figure}[htbp]
  \label{suppfig:diag-plot-mode}
  \centering
  \includegraphics[width=\textwidth]{figures/supplementary/phastcons_diag_plot.pdf}

  \caption{An optimization diagnostic plot showing an example of the bimodal likelihood
  optima, or a numeric optimization issue.}
\end{figure}

In one case, for our PhastCons Priority model mutation rate estimtes, our
jackknife standard error estimates were very large. Upon inspection of
optimization diagnostic plots (see Supplementary Materials Section
\ref{suppfig:diag-plot-mode}) and histograms of jackknife estimates, these
looked like they were due to rare aberrant optima. These were either caused by
a numeric optimization issue or another likelihood mode. Given that this looked
to be artificially inflating the uncertainty compared to all other models, we
filtered out these jackknife results before computing the standard error.

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[]{figures/supplementary/rescale_pi_pred_plot.pdf}

  \caption{Chromosome 2 predictions on the YRI samples, with locally-rescaled
      model fits. Observed data is the dark gray line, and the normal MLE for
      the PhastCons Priority model is the blue line. The locally-rescaled
      predictions are the green line. The dashed red line are the prediction
      using the standard B' map (without local-rescaled) and the maximum likelihood
      estimates from the locally-rescaled fits. The large discrepancy in this shows
      that estimates are highly dependent on the B' map.}
    \label{suppfig:rescale}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[]{figures/supplementary/rescale_Bs.pdf}

  \caption{The standard (dark gray) and locally-rescaled (blue) B' maps for different
      $\mu = 1.6 \times 10^{-8}$ and three different selection coefficients.
      For $s=10^{-5}$ ($2N_e s = 0.2$), locally-rescaling alters the predicted
      reduction so that it is essentially insignificant ($B \approx 1$). For
  mid-strength selection $s=10^{-4}$ ($2N_e s = 2$), there is only a very
slight difference between standard and locally-rescaled B' maps. Finally, for
strong selection $s=10^{-3}$ ($2N_e s = 20$), locally-rescaling does not change
the B' maps, as expected. }
    \label{suppfig:rescale-b}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[]{figures/supplementary/rescale_DFE.pdf}

@@

    \label{suppfig:rescale-dfe}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8\textwidth]{figures/supplementary/strongsel_grid_dfe.pdf}
  \caption{The DFE estimates for the strong selection grid (up to $s=10^{-1}$).}
    \label{suppfig:strong-sel-dfe}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8\textwidth]{figures/supplementary/ave_sel_vs_pi0.pdf}

@@

    \label{suppfig:sel-ident}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8\textwidth]{figures/supplementary/feature_priority_pop_dfe_strongselgrid.pdf}

  \caption{The DFE estimates for the strong selection grid (up to $s=10^{-1}$)
  under the Feature Priority model, across populations.}

    \label{suppfig:strong-sel-dfe-pop-phastcons}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8\textwidth]{figures/supplementary/phastcons_priority_pop_dfe_strongselgrid.pdf}

  \caption{The DFE estimates for the strong selection grid (up to $s=10^{-1}$)
  under the PhastCons Priority model, across populations.}
    \label{suppfig:strong-sel-dfe-pop-feature}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8\textwidth]{figures/supplementary/subfigure_sim_mle_expansion.pdf}

@@

    \label{suppfig:sim-exp}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.8\textwidth]{figures/supplementary/subfigure_sim_mle_recessive.pdf}

@@

  \label{suppfig:sim-rec}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6\textwidth]{figures/supplementary/cadd_residual.pdf}
  \caption{Residuals from the CADD 6\% sparse track plotted against the fraction of
  basepairs in a window annotated by a CADD region.}
  \label{suppfig:resid-cadd}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6\textwidth]{figures/supplementary/sweep_residuals.pdf}

@@

  \label{suppfig:resid-sweeps}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6\textwidth]{figures/supplementary/b_resid.pdf}
  \caption{Residuals from the CADD 6\% model plotted against the predicted $B(x)$
           value.}
  \label{suppfig:resid-b}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6\textwidth]{figures/supplementary/phastcons_resid_lof.pdf}
  \caption{Residuals from the PhastCons Priority model plotted against the
  average LoF selection coefficient across genes in a megabase window
(estimated from \cite{Agarwal2023-un}). These results show that the remaining likely 
selection signal is not dependent on the model used (PhastCons Priority versus CADD 6\%).}
  \label{suppfig:resid-pc-fit}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.9\textwidth]{figures/supplementary/figure_ceu_dfe.pdf}
  \caption{The DFE across models for the CEU reference sample (analogous to the
  DFE for YRI samples in Figure \ref{fig:figure-3}.}
  \label{suppfig:dfe-ceu}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.9\textwidth]{figures/supplementary/figure_chb_dfe.pdf}
  \caption{The DFE across models for the CHB reference sample (analogous to the
  DFE for YRI samples in Figure \ref{fig:figure-3}.}
    \label{suppfig:dfe-chb}
\end{figure}
